MAESTROeX: A Massively Parallel Low Mach Number Astrophysical Solver

### 2019-08-01 MAESTROeX_I/algorithm.tex
minor changes to sec 3 algorithm description

\item Compute the face-centered, time-centered species, $(\rho X_k)^{\nph,\pred}$,
  for the conservative update of $(\rho X_k)^{(1)}$ using a Godunov approach \citep{XRB_III}.
  As described in Paper V, for robust numerical slope limiting we predict 
  $\rho'^n=\rho^n-\rho_0^n$ and $X_k^n$ to faces
  and here we spatially interpolate $\rho_0^n$ to faces to assemble the fluxes.

\item Evolve $(\rho X_k)^{(1)} \rightarrow (\rho X_k)^{(2),\star}$ using
\begin{equation}
(\rho X_k)^{(2),\star} = (\rho X_k)^{(1)}
  - \dt \left\{ \nabla \cdot \left[ \uadvone (\rho X_k)^{\nph,\pred} \right] \right\},
\end{equation}

@@

\item Compute the face-centered, time-centered enthalpy, $(\rho h)^{\nph,\pred},$
  for the conservative update of $(\rho h)^{(1)}$ using using a Godunov approach \citep{XRB_III}.
  As described in Paper V, for robust numerical slope limiting 
  we predict $(\rho h)'^n=(\rho h)^n-(\rho h)_0^n$ to faces,
  where $(\rho h)_0^n$ is obtained by calling {\bf Average}$[(\rho h)^n]\rightarrow[(\rho h)_0]$,
  and here we spatially interpolate $(\rho h)_0^n$ to faces to assemble the fluxes.

\item Evolve $(\rho h)^{(1)} \rightarrow (\rho h)^{(2),\star}$ using
\begin{equation}
(\rho h)^{(2),\star}
= (\rho h)^{(1)} - \dt \left\{ \nabla \cdot \left[ \uadvone (\rho h)^{\nph,\pred} \right] \right\} + \Delta t\frac{Dp_0}{Dt}
\end{equation}
where here
\begin{equation}
\frac{Dp_0}{Dt} =
\begin{cases}
\frac{p_0^{n+1,*} - p_0^n}{\Delta t} + \left(\uadvone \cdot \eb_r\right) \left(\frac{\partial p_0}{\partial r} \right)^{n}& {\rm (planar)} \\
\frac{p_0^{n+1,*} - p_0^n}{\Delta t} + \left[ \nabla \cdot \left (\uadvone p_0^{\nph} \right ) - p_0^{\nph} \nabla \cdot \uadvone \right]& {\rm (spherical)}
\end{cases}
,
\end{equation}
and $p_0^\nph = (p_0^n+p_0^{n+1,*})/2$.
\end{enumerate}
\end{enumerate}

@@

\item Update $(\rho X_k)$.  This step is identical to {\bf Step 4A} except we use
  the updated values $\uadvtwo$ and $\rho_0^{n+1,\star}$ rather than
  $\uadvone$ and $\rho_0^{n+1,\star}$.  In particular:

@@

\item Compute the face-centered, time-centered species, $(\rho X_k)^{\nph,\pred}$,
  for the conservative update of $(\rho X_k)^{(1)}$ using a Godunov approach \citep{XRB_III}.
  Again, we predict $\rho'^n=\rho^n-\rho_0^n$ and $X_k^n$ to faces
  and here we spatially interpolate $(\rho_0^n+\rho_0^{n+1,*})/2$ to faces to assemble the fluxes.

@@

\item Update the enthalpy.  This step is identical to {\bf Step 4D} except we use
  the updated values $\uadvtwo$, $\rho_0^{n+1}$, $(\rho h)_0^{n+1}$, and $p_0^{\nph}$
  rather than
  $\uadvone, \rho_0^{n+1,\star}$, $(\rho h)_0^{n+1,\star}$, and $p_0^n$.
  In particular:

@@

\item Compute the face-centered, time-centered enthalpy, $(\rho h)^{\nph,\pred},$
  for the conservative update of $(\rho h)^{(1)}$ using a Godunov approach \citep{XRB_III}.
  Again, we predict $(\rho h)'^n=(\rho h)^n-(\rho h)_0^n$ to faces
  and here we spatially interpolate $[(\rho h)_0^n)+(\rho h)_0^{n+1,*}]/2$ to faces to assemble the fluxes.

@@

First, we compute the face-centered, time-centered velocities, $\Ub^{\nph,\pred}$
using a Godunov approach \citep{XRB_III}. Then, we define
\begin{equation}
\rho^\nph = \frac{\rho^n + \rho^{n+1}}{2}, \qquad \rho_0^\nph = \frac{\rho_0^n + \rho_0^{n+1}}{2}.
\end{equation}
We update the velocity field $\Ub^n$ to $\Ub^{n+1,\dagger}$ by discretizing
equation (\ref{eq:momentum}) as
\begin{equation}
\Ub^{n+1,\dagger}
= \Ub^n - \dt \left[\uadvtwo \cdot \nabla \Ub^{\nph,\pred} \right]
 - \dt \left[ \frac{\beta_0^\nph}{\rho^\nph} \nabla \left( \frac{\pi^\nmh}{\beta_0^\nmh} \right) + \frac{\left(\rho^\nph-\rho_0^\nph\right)}{\rho^\nph} g^{\nph} \eb_r \right],
\end{equation}
Again, the $\dagger$ superscript refers
to the fact that the updated velocity does not satisfy the divergence constraint,
\begin{equation}
\nabla \cdot \left(\beta_0^{n+1} \Ub^{n+1} \right) = \beta_0^{n+1} \left[ S^{n+1} - \frac{1}{\gammaonebar^{n+1}p_0^{n+1}}\left(\frac{\partial p_0}{\partial t}\right)^{\nph}\right].\label{eq:div3}
\end{equation}
We use an approximate projection to project $\Ub^{n+1,\dagger}$ onto the space of velocities that satisfy the constraint to obtain $\Ub^{n+1}$ using a ``nodal'' projection.
This projection necessarily differs from the MAC projection used in
{\bf Step 3} and {\bf Step 7} because the velocities in those steps are defined
on faces and $\Ub^{n+1}$ is defined at cell centers, requiring different divergence
and gradient operators.
Furthermore, as part of the nodal projection, we also define a nodal new-time perturbational pressure, $\pi^\nph$.
Refer to Appendix \ref{Sec:Projection} for more details.

>>>>

\item Compute the face-centered, time-centered species, $(\rho X_k)^{\nph,\pred,\star}$,
  for the conservative update of $(\rho X_k)^{(1)}$ using a Godunov approach \citep{XRB_III}.
  As described in Paper V, for robust numerical slope limiting we predict 
  $\rho'^n=\rho^n-\rho_0^n$ and $X_k^n$ to faces
  and here we spatially interpolate $\rho_0^n$ to faces to assemble the fluxes.

\item Evolve $(\rho X_k)^{(1)} \rightarrow (\rho X_k)^{(2),\star}$ using
\begin{equation}
(\rho X_k)^{(2),\star} = (\rho X_k)^{(1)}
  - \dt \left\{ \nabla \cdot \left[ \uadvone (\rho X_k)^{\nph,\pred,\star} \right] \right\},
\end{equation}

@@

\item Compute the face-centered, time-centered enthalpy, $(\rho h)^{\nph,\pred,\star},$
  for the conservative update of $(\rho h)^{(1)}$ using using a Godunov approach \citep{XRB_III}.
  As described in Paper V, for robust numerical slope limiting 
  we predict $(\rho h)'^n=(\rho h)^n-(\rho h)_0^n$ to faces,
  where $(\rho h)_0^n$ is obtained by calling {\bf Average}$[(\rho h)^n]\rightarrow[(\rho h)_0^n]$,
  and here we spatially interpolate $(\rho h)_0^n$ to faces to assemble the fluxes.

\item Evolve $(\rho h)^{(1)} \rightarrow (\rho h)^{(2),\star}$ using
\begin{equation}
(\rho h)^{(2),\star}
= (\rho h)^{(1)} - \dt \left\{ \nabla \cdot \left[ \uadvone (\rho h)^{\nph,\pred,\star} \right] \right\} + \Delta t\frac{Dp_0}{Dt}
\end{equation}
where here
\begin{equation}
\frac{Dp_0}{Dt} =
\begin{cases}
\frac{p_0^{n+1,*} - p_0^n}{\Delta t} + \left(\uadvone \cdot \eb_r\right) \left(\frac{\partial p_0}{\partial r} \right)^{n}& {\rm (planar)} \\
\frac{p_0^{n+1,*} - p_0^n}{\Delta t} + \left[ \nabla \cdot \left (\uadvone p_0^{\nph} \right ) - p_0^{\nph} \nabla \cdot \uadvone \right]& {\rm (spherical)}
\end{cases}
,
\end{equation}
and $p_0^\nph = (p_0^n+p_0^{n+1,*})/2$.
\end{enumerate}
\end{enumerate}

@@

\item Update $(\rho X_k)$.  This step is identical to {\bf Step 4A} except we use
  the updated values $\uadvtwo$ and $\rho_0^{n+1,\star}$ rather than
  $\uadvone$ and $\rho_0^n$.  In particular:

@@

\item Compute the face-centered, time-centered species, $(\rho X_k)^{\nph,\pred}$,
  for the conservative update of $(\rho X_k)^{(1)}$ using a Godunov approach \citep{XRB_III}.
  Again, we predict $\rho'^n=\rho^n-\rho_0^n$ and $X_k^n$ to faces
  but here we spatially interpolate $(\rho_0^n+\rho_0^{n+1,*})/2$ to faces to assemble the fluxes.

@@

\item Update the enthalpy.  This step is identical to {\bf Step 4D} except we use
  the updated values $\uadvtwo$, $\rho_0^{n+1}$, $(\rho h)_0^{n+1}$, and $p_0^{n+1}$
  rather than
  $\uadvone, \rho_0^n$, $(\rho h)_0^n$, and $p_0^n$.
  In particular:

@@

\item Compute the face-centered, time-centered enthalpy, $(\rho h)^{\nph,\pred},$
  for the conservative update of $(\rho h)^{(1)}$ using a Godunov approach \citep{XRB_III}.
  Again, we predict $(\rho h)'^n=(\rho h)^n-(\rho h)_0^n$ to faces
  but here we spatially interpolate $[(\rho h)_0^n)+(\rho h)_0^{n+1,*}]/2$ to faces to assemble the fluxes.

@@

First, we compute the face-centered, time-centered velocities, $\Ub^{\nph,\pred}$
using a Godunov approach \citep{XRB_III}. Then, we update
the velocity field $\Ub^n$ to $\Ub^{n+1,\dagger}$ by discretizing
equation (\ref{eq:momentum}) as
\begin{equation}
\Ub^{n+1,\dagger}
= \Ub^n - \dt \left[\uadvtwo \cdot \nabla \Ub^{\nph,\pred} \right]
 - \dt \left[ \frac{\beta_0^\nph}{\rho^\nph} \nabla \left( \frac{\pi^\nmh}{\beta_0^\nmh} \right) + \frac{\left(\rho^\nph-\rho_0^\nph\right)}{\rho^\nph} g^{\nph} \eb_r \right],
\end{equation}
where
\begin{equation}
\rho^\nph = \frac{\rho^n + \rho^{n+1}}{2}, \qquad \rho_0^\nph = \frac{\rho_0^n + \rho_0^{n+1}}{2}.
\end{equation}
Again, the $\dagger$ superscript refers
to the fact that the updated velocity does not satisfy the divergence constraint,
\begin{equation}
\nabla \cdot \left(\beta_0^{n+1} \Ub^{n+1} \right) = \beta_0^{n+1} \left[ S^{n+1} - \frac{1}{\gammaonebar^{n+1}p_0^{n+1}}\left(\frac{\partial p_0}{\partial t}\right)^{\nph}\right].\label{eq:div3}
\end{equation}
We use an approximate projection to project $\Ub^{n+1,\dagger}$ onto the space of velocities that satisfy the constraint to obtain $\Ub^{n+1}$ using a ``nodal'' projection.
This projection necessarily differs from the MAC projection used in
{\bf Step 3} and {\bf Step 7} because the velocities in those steps are defined
on faces and $\Ub^{n+1}$ is defined at cell centers, requiring different divergence
and gradient operators.
Furthermore, as part of the nodal projection, we also define a nodal new-time perturbational pressure, $\pi^\nph$.
Refer to Appendix \ref{Sec:Projection} for more details.
